---
title: "Combined Analysis"
output: html_document
---

# Set Up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 14)
```

```{r}
#clear workspace
rm(list=ls()) 
```

```{r}
#colour blind friendly colour palette for dataviz.
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
#custom functions

#get model output as latex
mod_out <- function(model_name) {
  # Ensure the model is of class 'lmerMod'
  class(model_name) <- "lmerMod"
  
  # Generate LaTeX output with stargazer
  stargazer(model_name,
            out = "model_output.tex", 
            star.cutoffs = c(0.05, 0.01, 0.001), 
            digits = 2, 
            no.space = TRUE, 
            omit.stat = "LL", 
            single.row = TRUE, 
            align = TRUE, 
            header = FALSE, 
            intercept.bottom = FALSE)
}

#output summary stats as latex
sum_tbl <- function(combined_latin) {
  summary_table <- combined_latin %>%
    group_by(Condition) %>%
    summarise(
      Mean_PP = mean(PP.RESID, na.rm = TRUE),
      SD_PP = sd(PP.RESID, na.rm = TRUE),
      Mean_VP = mean(VP.RESID, na.rm = TRUE),
      SD_VP = sd(VP.RESID, na.rm = TRUE),
      Mean_SPILL = mean(SPILL.RESID, na.rm = TRUE),
      SD_SPILL = sd(SPILL.RESID, na.rm = TRUE)
    ) %>%
    ungroup() %>%
    mutate(across(where(is.numeric), ~ round(.x, 2))) %>%
    kable(format = "latex", booktabs = TRUE,
          col.names = c("Condition", "Mean (ms)", "SD (ms)", 
                        "Mean (ms)", "SD (ms)", 
                        "Mean (ms)", "SD (ms)")) %>%
    kable_styling(latex_options = c("striped", "hold_position")) %>%
    add_header_above(c(" " = 1, "PP Region" = 2, "VP Region" = 2, "SPILL Region" = 2))
  
  cat(summary_table)
}
```


```{r, message=FALSE}
#load packages - nb: not all packages used
library(tidyverse)      #for tidying and wrangling
library(lme4)           #for LMM building
library(lmerTest)       #for better output from LMMs
library(emmeans)        #for our planned comparisons
library(performance)    #check model assumptions
library(buildmer)       #assists with iterating through covariates and ixns
library(ggpubr)         #plotting
library(texreg)         #for LaTeX table output
library(kableExtra)     #table formatting
library(xtable)         #table outputs
library(stargazer)      #lmm output
library(grid)           #multiple plots
library(patchwork)      #for plotting
library(stringr)        #string manipulation
library(scales)         #aesthetics
library(ggthemes)       #aesthetics
library(fitdistrplus)   #to check distribution
library(car)            #for vif() function
library(Hmisc)          #for rcorr() function
```

```{r}
#control scientific notation - makes reading p-values easier
options("scipen"=100000, "digits"=10)
#increase no. of calcs for emmeans functions
emm_options(pbkrtest.limit = 10000,
            lmerTest.limit = 10000)
```

# EXP2/3

```{r, message=FALSE}
#load datasets
exp2_tidied <- read_csv("exp2_tidied.csv")
exp3_tidied <- read_csv("exp3_tidied.csv")
```

## Wrangling

```{r}
#add experiment ID
exp3_tidied <- exp3_tidied %>%
  mutate(Experiment = "EXP3")
```

```{r}
#add experiment ID
exp2_tidied <- exp2_tidied %>%
  mutate(Experiment = "EXP2") %>%
  mutate(Participant = Participant + 59) #add 59 to participant number so we have unique participants
```

```{r}
#find columns in common
common_cols_latin <- Reduce(intersect, list(colnames(exp2_tidied), colnames(exp3_tidied)))
```

```{r}
#combine
combined_latin <- bind_rows(
  exp2_tidied %>% dplyr::select(all_of(common_cols_latin)),
  exp3_tidied %>% dplyr::select(all_of(common_cols_latin))
)
```

```{r}
#count participants - 119 (=60+59)
combined_latin %>%
  distinct(Participant) %>%
  count()
```

```{r}
#count conditions
combined_latin %>%
  count(Condition)
```

```{r}
#uneven number of obs - downsample
set.seed(76543) #ensure reproducibility
combined_latin <- combined_latin %>%
  group_by(Condition) %>%
  sample_n(885) %>%
  ungroup()
```

```{r}
#count conditions
combined_latin %>%
  count(Condition)
```

```{r}
#factorise
combined_latin <- combined_latin %>%
  mutate(RestrictorType = factor(RestrictorType, levels = c("OTH", "SUB", "MAX")),
       RestrictorPosition = factor(RestrictorPosition, levels = c("EXT", "INT")),
       Participant = factor(Participant),
       StimID = factor(StimID))
```

```{r}
#contrast code
contrasts(combined_latin$RestrictorPosition) <- matrix(c(.5, -.5))
contrasts(combined_latin$RestrictorType) <- contr.sum(3)
```

## Descriptives & dataviz for RTs for each chunk

```{r}
#output summary stats
sum_tbl(combined_latin)
```

```{r, message=FALSE, warning=FALSE}
#pull out chunk names
roi_latin <- grep("^C\\d+\\.RT.RESID$|^SPILL\\.RT.RESID$", names(combined_latin), value = TRUE)

chunks_latin <- combined_latin %>%
  pivot_longer(cols = all_of(roi_latin), 
               names_to = "Chunk", 
               values_to = "ReadingTime") %>%
  group_by(Condition, Chunk) %>%
  summarise(
    mean_RT = mean(ReadingTime),
    sd_RT = sd(ReadingTime),
    se_RT = sd_RT / sqrt(n())
  ) %>%
  separate(Condition, into = c("RestrictorPosition", "RestrictorType")) %>%
  mutate(ChunkLabel = case_when(
    RestrictorPosition == "EXT" & Chunk == "C1.RT.RESID" ~ "EVERY",
    RestrictorPosition == "EXT" & Chunk == "C2.RT.RESID" ~ "NP",
    RestrictorPosition == "EXT" & Chunk == "C3.RT.RESID" ~ "VP",
    RestrictorPosition == "EXT" & Chunk == "C4.RT.RESID" ~ "PP",
    RestrictorPosition == "EXT" & Chunk == "SPILL.RT.RESID" ~ "SPILL",
    RestrictorPosition == "INT" & Chunk == "C1.RT.RESID" ~ "EVERY",
    RestrictorPosition == "INT" & Chunk == "C2.RT.RESID" ~ "NP",
    RestrictorPosition == "INT" & Chunk == "C3.RT.RESID" ~ "PP",
    RestrictorPosition == "INT" & Chunk == "C4.RT.RESID" ~ "VP",
    RestrictorPosition == "INT" & Chunk == "SPILL.RT.RESID" ~ "SPILL",
    TRUE ~ Chunk
  ))

```

```{r}
#plots
latin_ext <- chunks_latin %>%
  filter(RestrictorPosition == "EXT") %>%
  ggplot(aes(x = Chunk, y = mean_RT, group = RestrictorType, color = RestrictorType)) +
  geom_ribbon(aes(ymin = mean_RT - se_RT, ymax = mean_RT + se_RT, group = RestrictorType), 
              fill = "grey", alpha = 0.2, linetype = "blank") +
  geom_line(aes(linetype = RestrictorType), linewidth = 1) +  
  geom_point(aes(shape = RestrictorType), size = 3) +  
  scale_y_continuous(name = "Mean Reading Time (ms.)") +
  theme_light(base_size = 20, base_family = "Helvetica") +
  scale_colour_manual(values = cbPalette) +  
  facet_grid(vars(RestrictorPosition), switch = "y") +
  theme(legend.background = element_rect(color = "black", linewidth = 1),
        plot.title = element_text(hjust = 0.5, size = 25),
        strip.background = element_rect(fill = "grey90"),
        strip.text = element_text(colour = "black")) +
  scale_x_discrete(labels = c("EVERY", "NP", "VP", "PP", "SPILL"))

latin_int <- chunks_latin %>%
  filter(RestrictorPosition == "INT") %>%
  ggplot(aes(x = Chunk, y = mean_RT, group = RestrictorType, color = RestrictorType)) +
  geom_ribbon(aes(ymin = mean_RT - se_RT, ymax = mean_RT + se_RT, group = RestrictorType), 
              fill = "grey", alpha = 0.2, linetype = "blank") +
  geom_line(aes(linetype = RestrictorType), linewidth = 1) +  
  geom_point(aes(shape = RestrictorType), size = 3) +  
  scale_y_continuous(name = "Mean Reading Time (ms.)") +
  theme_light(base_size = 20, base_family = "Helvetica") +
  scale_colour_manual(values = cbPalette) +  
  facet_grid(vars(RestrictorPosition), switch = "y") +
  theme(legend.background = element_rect(color = "black", linewidth = 1),
        plot.title = element_text(hjust = 0.5, size = 25),
        strip.background = element_rect(fill = "grey90"),
        strip.text = element_text(colour = "black")) +
  scale_x_discrete(labels = c("EVERY", "NP", "PP", "VP", "SPILL"))
```

```{r}
#combine plots
pdf("charts/COMB_LATIN_chunks.pdf", width = 14, height = 14)
latin_ext + 
  latin_int + 
  plot_layout(ncol = 1,
              guides = "collect",
              axis_titles = "collect") +
  plot_annotation(title = "Mean Chunk Reading Times (ms.)",
                  theme = theme(plot.title = element_text(hjust = 0.5,
                                                          size = 30)))
```

## Reading Time Analysis for Preposition Phrase

```{r}
#dataviz for condition - logged for better viz.
pdf("charts/COMB_LATIN_pp_plot.pdf", width = 14, height = 14)
pp_ltn <- combined_latin %>%
  ggplot(aes(x = RestrictorType, #group by condition on x axis
             y = PP.RESID.LOG, #log RTs on y axis
             fill = RestrictorType)) + #colour by condition
  theme_light(base_size = 25, #default font size
              base_line_size = .5, #default line weight
              base_rect_size = 1, #default polygon border weight
              base_family = "Helvetica") + #font family Helvetica
  geom_violin(trim = FALSE) + #violin, don't trim ends
  geom_boxplot(width = 0.3, #boxplot
               outlier.shape = NA) + #don't show outlier dots
  scale_fill_manual(values = cbPalette) + #colour blind friendly palette
  labs(title = "PP Reading Times (ms.)",
       y = "Logged Reading Time (ms.)",
       x = NULL) +
  theme(legend.background = element_rect(color = "black", linewidth = 1), #box around legend
        plot.title = element_text(hjust = 0.5, size = 30), #center title, title font size
        legend.position = "bottom") + #legend at bottom
  coord_flip() + #flip coordinates so it's horizontal
  facet_wrap(~RestrictorPosition) +
  theme(legend.background = element_rect(color = "black", linewidth = 1),
        plot.title = element_text(hjust = 0.5, size = 25),
        strip.background = element_rect(fill = "grey90"),
        strip.text = element_text(colour = "black"))
pp_ltn
```

### Modelling

```{r, results=FALSE, message=FALSE}
#building a buildmer model
pp.ltn.blmr1 <- buildmer(PP.RESID ~ RestrictorType * RestrictorPosition + 
                       (1 + RestrictorType * RestrictorPosition | Participant) +
                       (1 + RestrictorType * RestrictorPosition | StimID),
                     buildmerControl = buildmerControl(include = ~ RestrictorType * RestrictorPosition),
                     data = combined_latin)
```

```{r}
#model summary
summary(pp.ltn.blmr1)
```

```{r}
#increase ranfx
pp.ltn1 <- lmer(PP.RESID ~ RestrictorType * RestrictorPosition +
                  (1 + RestrictorPosition | Participant) +
                  (1 | StimID),
                data = combined_latin)
summary(pp.ltn1) #OK, main fx
```

```{r}
#increase ranfx complexity
pp.ltn2 <- lmer(PP.RESID ~ RestrictorType * RestrictorPosition +
                  (1 + RestrictorPosition | Participant) +
                  (1 + RestrictorPosition | StimID),
                data = combined_latin)
summary(pp.ltn2) #singular fit
```

```{r}
#increase ranfx complexity
pp.ltn3 <- lmer(PP.RESID ~ RestrictorType * RestrictorPosition +
                  (1 + RestrictorType + RestrictorPosition | Participant) +
                  (1 | StimID),
                data = combined_latin)
summary(pp.ltn3) #singular fit, convergence issue
```

```{r}
#check model
check_model(pp.ltn1) #OK
```

```{r}
#add covars
pp.ltn4 <- lmer(PP.RESID ~ RestrictorType * RestrictorPosition +
                  TrialN + Experiment +
                  (1 + RestrictorPosition | Participant) +
                  (1 | StimID),
                data = combined_latin)
summary(pp.ltn4) #ok, main fx, ixn
```

```{r}
check_model(pp.ltn4) #ok
```

```{r}
#compare - covar mod better
anova(pp.ltn1, pp.ltn4)
```

```{r}
#model output
mod_out(pp.ltn4)
```

```{r}
#emms
pp.ltn_emms1 <- emmeans(pp.ltn4, ~ RestrictorType * RestrictorPosition,
                   adjust = "none") #we'll adjust when doing contrasts
```

```{r}
#contrasts
pp.ltn_contrasts1 <- contrast(pp.ltn_emms1,
                 adjust = "tukey",
                 "pairwise", 
                 simple = "RestrictorType")
```

```{r}
#print latex
pp.ltn_contrasts1 %>%
  xtable() 
#early fx could be priming, later fx less so.
#in future work, I'll try and devise
#a better paradigm to mitigate against this
```

#### TrialN-Mediated Interaction?

```{r}
#exp3 showed an ixn mediated by TrialN, so let's look at that
pp_ltn_int_trials <- combined_latin %>%
  filter(RestrictorPosition == "INT") %>%
  ggplot(aes(x = TrialN, y = PP.RESID, color = RestrictorType, linetype = RestrictorType)) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.1) +
  labs(x = NULL, y = "Reading Time") +
  theme_light(base_size = 25) +
  theme(plot.title = element_text(size = 30, hjust = 0.5),
        legend.background = element_rect(color = "black", linewidth = 1),
        legend.position = "bottom") +
  facet_grid(RestrictorPosition~Experiment) +
  scale_colour_manual(values = cbPalette)

pp_ltn_ext_trials <- combined_latin %>%
  filter(RestrictorPosition == "EXT") %>%
  ggplot(aes(x = TrialN, y = PP.RESID, color = RestrictorType, linetype = RestrictorType)) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.1) +
  labs(x = "Trial Number", y = "Reading Time") +
  theme_light(base_size = 25) +
  theme(plot.title = element_text(size = 30, hjust = 0.5),
        legend.background = element_rect(color = "black", linewidth = 1),
        legend.position = "bottom") +
  facet_grid(RestrictorPosition~Experiment) +
  scale_colour_manual(values = cbPalette)
```

```{r}
#combine plots - looks like different adaptation profiles across experiments
pdf("charts/COMB_LATIN_trials_plot.pdf", width = 14, height = 14)
pp_ltn_int_trials + 
  pp_ltn_ext_trials + 
  plot_layout(ncol = 1,
              guides = "collect",
              axis_titles = "collect") +
  plot_annotation(title = "PP Reading Times (ms.) by Trial x Condition",
                  theme = theme(plot.title = element_text(hjust = 0.5,
                                                          size = 30),
                                legend.background = element_rect(color = "black", linewidth = 1),
                                legend.position = "bottom"))
```

```{r}
#model
pp.ltn5 <- lmer(PP.RESID ~ RestrictorType * RestrictorPosition * scale(TrialN) + 
                  Experiment +
                  (1 + RestrictorPosition | Participant) +
                  (1 | StimID),
                data = combined_latin)
summary(pp.ltn5) #ok, main fx, ixn
```

```{r}
#check
check_model(pp.ltn5) #ok
```

```{r}
#mod out
mod_out(pp.ltn5)
```

```{r}
#generate sequence of trials
trial_seq <- seq(min(combined_latin$TrialN), max(combined_latin$TrialN), length.out = 100)

#predict RTs for levels of TrialN given ixn
predicted_data <- emmeans(pp.ltn5, ~ RestrictorType * RestrictorPosition * scale(TrialN), 
                          at = list(TrialN = trial_seq), type = "response")

#convert to df for dataviz
pred_df <- as.data.frame(predicted_data)
```

```{r}
#plot the predicted RTs over trials - adaptation rate is different; EXT-OTH and INT-OTH get easier, though at different rates
pdf("charts/COMB_LATIN_trials_preds_plot.pdf", width = 14, height = 14)
ltn_trials_preds <- pred_df %>%
  ggplot(aes(x = TrialN, y = emmean, color = RestrictorType, linetype = RestrictorType)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = emmean - SE, ymax = emmean + SE, fill = RestrictorType), alpha = 0.2) +
  scale_colour_manual(values = cbPalette) +
  scale_fill_manual(values = cbPalette) +
  labs(x = "TrialN", y = "Reading Time") +
  facet_grid(~RestrictorPosition) +
  theme_light(base_size = 25) +
  theme(plot.title = element_text(size = 30, hjust = 0.5),
        legend.background = element_rect(color = "black", linewidth = 1),
        legend.position = "bottom")
ltn_trials_preds
```

```{r}
#emms
pp.ltn_emms2 <- emmeans(pp.ltn5, ~ RestrictorType * RestrictorPosition,
                   adjust = "none") #we'll adjust when doing contrasts
```

```{r}
#contrasts
pp.ltn_contrasts2 <- contrast(pp.ltn_emms2,
                 adjust = "tukey",
                 "pairwise", 
                 simple = "RestrictorType")
```

```{r}
#print latex table
pp.ltn_contrasts2 %>%
  xtable()
```

```{r}
#plot it
pdf("charts/COMB_LATIN_emms_ixn_plot.pdf", width = 14, height = 14)
ltn_pp_ixn_plot <- emmip(pp.ltn5, 
                            RestrictorType ~ RestrictorPosition,
                           CIs = TRUE,
                           CIarg = list(lwd = .5,
                                        alpha = 1)) +
  scale_colour_manual(values = cbPalette) + #colour blind friendly palette
  labs(title = "PP EMM Reading Times Interaction Plot",
       y = "Estimated Marginal Mean Reading Time (ms.)",
       x = "RestrictorPosition") + #set up labels/title
  theme_light(base_size = 25, #default font size
              base_line_size = .5, #default line weight
              base_rect_size = 1, #default polygon border weight
              base_family = "Helvetica") + #font family Helvetica
  theme(plot.title = element_text(size = 30, #title font size
                                hjust = 0.5), #center title
      legend.background = element_rect(color = "black", linewidth = 1), #box around legend
      legend.position = "bottom") #legend at bottom
ltn_pp_ixn_plot
```

### Could there be priming effects?

```{r}
#add cumulative reading time
combined_latin <- combined_latin %>%
  mutate(CumulativeRT = case_when(RestrictorPosition == "INT" ~ C1.RT + C2.RT,
                                  RestrictorPosition == "EXT" ~ C1.RT + C2.RT + C3.RT)) %>%
  mutate(CumulativeRT = scale(CumulativeRT))
```

```{r}
ltn_cumulative_plot_int <- combined_latin %>%
  filter(RestrictorPosition == "INT") %>%
  ggplot(aes(x = CumulativeRT, y = PP.RESID, color = RestrictorType, linetype = RestrictorType)) +
  geom_smooth(method = "gam", se = TRUE, alpha = 0.1) + #gam smoothing b/c it takes into account non-linear trends
  labs(x = "Cumulative Reading Time", y = "PP Reading Time") +
  theme_light(base_size = 25) +
  facet_wrap(~RestrictorPosition, switch = "y") +
  theme(plot.title = element_text(size = 30, hjust = 0.5),
        legend.background = element_rect(color = "black", linewidth = 1),
        legend.position = "bottom") +
  scale_colour_manual(values = cbPalette)
```

```{r}
ltn_cumulative_plot_ext <- combined_latin %>%
  filter(RestrictorPosition == "EXT") %>%
  ggplot(aes(x = CumulativeRT, y = PP.RESID, color = RestrictorType, linetype = RestrictorType)) +
  geom_smooth(method = "gam", se = TRUE, alpha = 0.1) + #gam smoothing b/c it takes into account non-linear trends
  labs(x = "Cumulative Reading Time", y = "PP Reading Time") +
  theme_light(base_size = 25) +
  facet_wrap(~RestrictorPosition, switch = "y") +
  theme(plot.title = element_text(size = 30, hjust = 0.5),
        legend.background = element_rect(color = "black", linewidth = 1),
        legend.position = "bottom") +
  scale_colour_manual(values = cbPalette)
```


```{r}
#combine plots - looks like different adaptation profiles across experiments
pdf("charts/COMB_LATIN_cumulativeRT_effect.pdf", width = 14, height = 14)
ltn_cumulative_plot_int + 
  ltn_cumulative_plot_ext + 
  plot_layout(ncol = 1,
              guides = "collect",
              axis_titles = "collect") +
  plot_annotation(title = "Effect of Cumulative Reading Time on PP Reading Time",
                  theme = theme(plot.title = element_text(hjust = 0.5,
                                                          size = 30),
                                legend.background = element_rect(color = "black", linewidth = 1),
                                legend.position = "bottom")) #can only be taken tentatively, but effect reduces in INT-OTH condition
```


```{r}
#model
pp.ltn6 <- lmer(PP.RESID ~ RestrictorType * RestrictorPosition * TrialN * CumulativeRT + 
                  Experiment +
                  (1 + RestrictorPosition | Participant) +
                  (1 | StimID),
                data = combined_latin)
summary(pp.ltn6) #ok, main fx, ixn
```

```{r}
mod_out(pp.ltn6)
```

## Reading Time Analysis for Verb Phrase

```{r}
#dataviz for condition - logged for better viz.
pdf("charts/COMB_LATIN_vp_plot.pdf", height = 14, width = 14)
vp_viz <- combined_latin %>%
  ggplot(aes(x = RestrictorType, #group by condition on x axis
             y = VP.RESID.LOG, #log RTs on y axis
             fill = RestrictorType)) + #colour by condition
  theme_light(base_size = 25, #default font size
              base_line_size = .5, #default line weight
              base_rect_size = 1, #default polygon border weight
              base_family = "Helvetica") + #font family Helvetica
  geom_violin(trim = FALSE) + #violin, don't trim ends
  geom_boxplot(width = 0.3, #boxplot
               outlier.shape = NA) + #don't show outlier dots
  scale_fill_manual(values = cbPalette) + #colour blind friendly palette
  labs(title = "VP Reading Times (ms.)",
       y = "Logged Reading Time (ms.)",
       x = NULL) +
  theme(legend.background = element_rect(color = "black", linewidth = 1), #box around legend
        plot.title = element_text(hjust = 0.5, size = 30), #center title, title font size
        legend.position = "bottom") + #legend at bottom
  coord_flip() + #flip coordinates so it's horizontal
  facet_wrap(~RestrictorPosition) +
  theme(legend.background = element_rect(color = "black", linewidth = 1),
        plot.title = element_text(hjust = 0.5, size = 25),
        strip.background = element_rect(fill = "grey90"),
        strip.text = element_text(colour = "black"))
#print chart
vp_viz
```

### Modelling

```{r}
#buildmer model
vp.ltn.blmr <- buildmer(VP.RESID ~ RestrictorType * RestrictorPosition +
                       (1 + RestrictorType * RestrictorPosition | Participant) +
                       (1 + RestrictorType * RestrictorPosition | StimID),
                       data = combined_latin)
```

```{r}
summary(vp.ltn.blmr) #main fx, ixn
```

```{r}
#manual model
vp.ltn1 <- lmer(VP.RESID ~ RestrictorType * RestrictorPosition +
                  (1 + RestrictorPosition | Participant) +
                  (1 | StimID),
                data = combined_latin)
summary(vp.ltn1) #main fx, ixn
```

```{r}
#increase ranfx complexity
vp.ltn2 <- lmer(VP.RESID ~ RestrictorType * RestrictorPosition +
                  (1 + RestrictorPosition | Participant) +
                  (1 + RestrictorType | StimID),
                data = combined_latin)
summary(vp.ltn2) #singular fit
```

```{r}
#increase ranfx complexity
vp.ltn3 <- lmer(VP.RESID ~ RestrictorType * RestrictorPosition +
                  (1 + RestrictorPosition | Participant) +
                  (1 + RestrictorPosition | StimID),
                data = combined_latin)
summary(vp.ltn3) #singular fit
```

```{r}
#add covars
vp.ltn4 <- lmer(VP.RESID ~ RestrictorType * RestrictorPosition +
                  TrialN +
                  (1 + RestrictorPosition | Participant) +
                  (1 | StimID),
                data = combined_latin)
summary(vp.ltn4) #main fx, ixn
```

```{r}
#check model
check_model(vp.ltn4) #ok
```

```{r}
#compare - covar mod better
anova(vp.ltn1, vp.ltn4)
```

```{r}
#model output
mod_out(vp.ltn4)
```

```{r}
#emms
vp.ltn_emms1 <- emmeans(vp.ltn4, ~ RestrictorType * RestrictorPosition,
                   adjust = "none")
vp.ltn_emms1
```

```{r}
#just look at INT
vp.ltn.int_emms1 <- subset(vp.ltn_emms1, RestrictorPosition == "INT")
```

```{r}
#contrasts
contrast(vp.ltn.int_emms1,
         adjust = "tukey",
         "pairwise") %>%
  xtable()
```

```{r}
#checking that INT conditions are slower that EXT conditions overall
vp.ltn.pos.emms1 <- emmeans(vp.ltn4,
                       pairwise ~ RestrictorPosition,
                       adjust = "none") #only one comparison made
vp.ltn.pos.emms1
```

```{r}
#plot it
pdf("charts/COMB_LATIN_vp_ixn_plot.pdf", height = 14, width = 14)
vp_ixn_plot <- emmip(vp.ltn4,
                     RestrictorType ~ RestrictorPosition,
                     CIs = TRUE,
                     CIarg = list(lwd = .5,
                                  alpha = 1)) +
  scale_colour_manual(values = cbPalette) + #colour blind friendly palette
  labs(title = "VP EMM Reading Times Interaction Plot",
       y = "Estimated Marginal Mean Reading Time (ms.)",
       x = "RestrictorPosition") + #set up labels/title
  theme_light(base_size = 25, #default font size
              base_line_size = .5, #default line weight
              base_rect_size = 1, #default polygon border weight
              base_family = "Helvetica") + #font family Helvetica
  theme(plot.title = element_text(size = 30, #title font size
                                  hjust = 0.5), #center title
        legend.background = element_rect(color = "black", linewidth = 1), #box around legend
        legend.position = "bottom") #legend at bottom
vp_ixn_plot
```

## Dataviz Showing Adjusted Means For ROIs

```{r}
#extract emms
vp_tbl <- tibble(as.data.frame(vp.ltn_emms1))
pp_tbl <- tibble(as.data.frame(pp.ltn_emms2))
```

```{r}
#dataviz showing raw mean RTs for quant and N regions
#and EMMs for PP, VP and SPILL regions

#list all emm_list objects
all_latin_emms <- list(
  pp_tbl %>% #PP emms from modelling
    mutate(Chunk = case_when(RestrictorPosition == "INT" ~ "C3.RT.RESID",
                             RestrictorPosition == "EXT" ~ "C4.RT.RESID")),
  vp_tbl %>% #VP emms from modelling
    mutate(Chunk = case_when(RestrictorPosition == "INT" ~ "C4.RT.RESID",
                             RestrictorPosition == "EXT" ~ "C3.RT.RESID"))) %>%
  map_dfr(as_tibble) %>% #create tibble from list
  rename(mean_RT = emmean, #match names of mean and SE in chunks_data
         se_RT = SE) %>%
  mutate(Condition = str_c(RestrictorPosition, "-", RestrictorType)) %>% #add Condition column
  mutate(mean_Type = "EMM") #add mean_Type column for reference
  
  
#extract raw means for quant and noun from chunks_data
raw_latin_means <- chunks_latin %>% 
  filter(Chunk %in% c("C1.RT.RESID", "C2.RT.RESID", "SPILL.RT.RESID")) %>% #filter chunks of interest
  mutate(mean_Type = "RAW") #add mean_Type column for reference

#bind to produce tibble
chunks_latin_adj <- bind_rows(raw_latin_means, all_latin_emms) %>%
  mutate(ChunkLabel = case_when(
    RestrictorPosition == "EXT" & Chunk == "C1.RT.RESID" ~ "EVERY",
    RestrictorPosition == "EXT" & Chunk == "C2.RT.RESID" ~ "NP",
    RestrictorPosition == "EXT" & Chunk == "C3.RT.RESID" ~ "VP",
    RestrictorPosition == "EXT" & Chunk == "C4.RT.RESID" ~ "PP",
    RestrictorPosition == "EXT" & Chunk == "SPILL.RT.RESID" ~ "SPILL",
    RestrictorPosition == "INT" & Chunk == "C1.RT.RESID" ~ "EVERY",
    RestrictorPosition == "INT" & Chunk == "C2.RT.RESID" ~ "NP",
    RestrictorPosition == "INT" & Chunk == "C3.RT.RESID" ~ "PP",
    RestrictorPosition == "INT" & Chunk == "C4.RT.RESID" ~ "VP",
    Chunk == "SPILL.RT.RESID" ~ "SPILL",
    TRUE ~ Chunk  # Default case, if no specific condition
  ))

```

```{r}
#plots
ext_ltn_adj <- chunks_latin_adj %>%
  mutate(RestrictorPosition = replace_na(RestrictorPosition, "EXT")) %>%
  filter(RestrictorPosition == "EXT") %>%
  ggplot(aes(x = Chunk, y = mean_RT, group = RestrictorType, color = RestrictorType)) +
  geom_ribbon(aes(ymin = mean_RT - se_RT, ymax = mean_RT + se_RT, group = RestrictorType), 
              fill = "grey", alpha = 0.2, linetype = "blank") +
  geom_line(aes(linetype = RestrictorType), linewidth = 1) +  
  geom_point(aes(shape = RestrictorType), size = 3) +  
  scale_y_continuous(name = "Mean Reading Time (ms.)") +
  theme_light(base_size = 20, base_family = "Helvetica") +
  scale_colour_manual(values = cbPalette) +  
  facet_grid(vars(RestrictorPosition), switch = "y") +
  theme(legend.background = element_rect(color = "black", linewidth = 1),
        plot.title = element_text(hjust = 0.5, size = 25),
        strip.background = element_rect(fill = "grey90"),
        strip.text = element_text(colour = "black")) +
  scale_x_discrete(labels = c("EVERY", "NP", "VP", "PP", "SPILL"))

int_ltn_adj <- chunks_latin_adj %>%
  mutate(RestrictorPosition = replace_na(RestrictorPosition, "INT")) %>%
  filter(RestrictorPosition == "INT") %>%
  ggplot(aes(x = Chunk, y = mean_RT, group = RestrictorType, color = RestrictorType)) +
  geom_ribbon(aes(ymin = mean_RT - se_RT, ymax = mean_RT + se_RT, group = RestrictorType), 
              fill = "grey", alpha = 0.2, linetype = "blank") +
  geom_line(aes(linetype = RestrictorType), linewidth = 1) +  
  geom_point(aes(shape = RestrictorType), size = 3) +  
  scale_y_continuous(name = "Mean Reading Time (ms.)") +
  theme_light(base_size = 20, base_family = "Helvetica") +
  scale_colour_manual(values = cbPalette) +  
  facet_grid(vars(RestrictorPosition), switch = "y") +
  theme(legend.background = element_rect(color = "black", linewidth = 1),
        plot.title = element_text(hjust = 0.5, size = 25),
        strip.background = element_rect(fill = "grey90"),
        strip.text = element_text(colour = "black")) +
  scale_x_discrete(labels = c("EVERY", "NP", "PP", "VP", "SPILL"))
```

```{r}
#combine plots
pdf("charts/COMB_LATIN_chunks_adj_plot.pdf", width = 14, height = 14)
ext_ltn_adj + 
  int_ltn_adj + 
  plot_layout(ncol = 1,
              guides = "collect",
              axis_titles = "collect") +
  plot_annotation(title = "Adjusted Reading Times (ms.)",
                  theme = theme(plot.title = element_text(hjust = 0.5,
                                                          size = 30)))
```

## Quick check on LOOKED vs. SEEMED

```{r}
percept.ltn1 <- lm(PP.RESID ~ LOOK_SEEM, data = combined_latin)
summary(percept.ltn1) #nothing

percept.ltn2 <- lm(VP.RESID ~ LOOK_SEEM, data = combined_latin)
summary(percept.ltn2) #nothing
```

# EXP1/2 INT-SUB vs. INT-MAX

## Wrangling

```{r}
#load exp1 dataset
exp1_tidied <- read_csv("exp1_tidied.csv") %>%
  mutate(Experiment = "EXP1") %>%
  filter(Condition != "NIL") %>%
  mutate(has_comp = as.character(has_comp))
```
```{r}
exp2_COPY_tidied <- exp2_tidied %>%
  mutate(has_comp = as.character(has_comp)) %>% #change has_comp to dbl
  mutate(Participant = Participant + 1) #add 1 to avoid overlap
```

```{r}
#find columns in common
common_cols_int <- Reduce(intersect, list(colnames(exp1_tidied), colnames(exp2_COPY_tidied)))
```

```{r}
#combine
combined_int <- bind_rows(
  exp1_tidied %>% dplyr::select(all_of(common_cols_int)),
  exp2_COPY_tidied %>%
    filter(RestrictorPosition == "INT")
  %>% dplyr::select(all_of(common_cols_int)))
```

```{r}
#count participants - 120
combined_int %>%
  distinct(Participant) %>%
  count()
```
```{r}
#ensure Condition is labelled consistently
combined_int <- combined_int %>%
  mutate(Condition = RestrictorType)
```


```{r}
#add Total Sentence Reading time raw and predicted
combined_int <- combined_int %>%
  rowwise() %>%
  mutate(TotalTime = sum(C1.RT, C2.RT, C3.RT, C4.RT, SPILL.RT),
         TotalPRED = sum(C1.PRED, C2.PRED, C3.PRED, C4.PRED, SPILL.PRED)) %>%
  ungroup()
```

## Descriptives & dataviz for RTs for each chunk

```{r, message=FALSE, warning=FALSE}
#pull out chunk names
roi_int <- grep("^C\\d+\\.RT.RESID$|^SPILL\\.RT.RESID$", names(combined_int), value = TRUE)

chunks_int <- combined_int %>%
  pivot_longer(cols = all_of(roi_int), 
               names_to = "Chunk", 
               values_to = "ReadingTime") %>%
  group_by(RestrictorType, Chunk, Experiment) %>%
  summarise(
    mean_RT = mean(ReadingTime),
    sd_RT = sd(ReadingTime),
    se_RT = sd_RT / sqrt(n())
  )
```

```{r}
#plots
pdf("charts/COMB_INT_chunks_plot.pdf", width = 14, height = 14)
combined_int_chunks_plot <- chunks_int %>%
  ggplot(aes(x = Chunk, y = mean_RT, group = RestrictorType, color = RestrictorType)) +
  geom_ribbon(aes(ymin = mean_RT - se_RT, ymax = mean_RT + se_RT, group = RestrictorType), 
              fill = "grey", alpha = 0.2, linetype = "blank") +
  geom_line(aes(linetype = RestrictorType), linewidth = 1) +
  geom_point(aes(shape = RestrictorType), size = 3) +
  scale_y_continuous(name = "Mean Reading Time (ms.)") +
  theme_light(base_size = 20, base_family = "Helvetica") +
  scale_colour_manual(values = cbPalette) +
  facet_grid(~ Experiment) +
  theme(legend.background = element_rect(color = "black", linewidth = 1),
        plot.title = element_text(hjust = 0.5, size = 25),
        strip.background = element_rect(fill = "grey90"),
        strip.text = element_text(colour = "black")) +
  scale_x_discrete(labels = c("EVERY", "NP", "PP", "VP", "SPILL"))
combined_int_chunks_plot
```

```{r}
#output summary stats exp1
combined_int %>%
  filter(Experiment == "EXP1") %>%
  sum_tbl()
```

```{r}
#output summary stats exp2
combined_int %>%
  filter(Experiment == "EXP2") %>%
  sum_tbl()
```

Look at whether reading times on the whole are slower in Exp1, suggesting an experiment-specific reading strategy. Look at: total sentence reading time, PP reading time. Also look at quantifying effect sizes if poss just to see if Exp1 effects were `real`.

Also look at whether predicted reading times are on the whole slower, because those predicted reading times were independent of the experimental trials -- they came from filler data. If both actual reading times and predicted reading times are slower in Exp1 compared to Exp2, that would suggest that participants in Exp1 were simply being much more conscientious/attentive readers, meaning that when the PP was presented, they were more aware of the interpretation that it would give rise to.

REMEMBER: look at first vs last XX% of trials. If you get the same effects in early trials in both Exp1 and Exp2, then it's not just the comprehension questions -- the effects arose due to the design of the experiment; e.g., because participants learned NOT to commit early in Exp2 and to wait to see if they got any more info later.

There are then two plausible interpretations for the differences. First, maybe Exp1's early findings actually **DO** suggest that participants were committing early. Second, maybe Exp1's findings indicate instead that there are expectations about likely restrictions, and once people are reading for comprehension they are more aware of the range of possibilities and aware of their preference for how the QNP `should' be restricted. Further work will need to manipulate the proportion of comprehension questions to determine whether that has a reliable influence on the effects. If the default is to commit early, reducing the proportion of comprehension questions should mean that early effects diminish as a function of the reduction. Given the findings from Experiment Four, I think Experiment One's findings can at least tentatively be interpreted as indicating expectation or anticipation. It would be too premature to claim that enforcing deeper comprehension FORCES commitment at an earlier stage in processing. The VP has been hypothesised to be a processing domain for scope processing, for instance. And these people find evidence for delayed aspectual processing: https://pubmed.ncbi.nlm.nih.gov/32672991/. See here for The Minimal Semantic Commitment Hypothesis: https://www.sciencedirect.com/science/article/abs/pii/S0010027799000025. See also: https://www.sciencedirect.com/science/article/abs/pii/0749596X90900717. Maybe of interest: http://act-r.psy.cmu.edu/wordpress/wp-content/uploads/2012/12/399sdarticle-11.pdf. On delay, look at Beck and Tiemann paper again (`englightened incrementality'). Arregui showed that quantifiers `look for' their restrictor and scope arguments. So expectation/anticipation effects are not totally unbelievable. Also look at:

https://www.semanticscholar.org/paper/Parameters-in-the-theory-of-sentence-processing%3A-Weinberg/1bd8a41ead3bae02b69be1b509edd9c9792b805e

https://philarchive.org/archive/VICSM

https://semanticsarchive.net/Archive/Tc2M2ZlM/The%20Logicality%20of%20language%20and%20Contextualism%20vs.%20Semantic%20Minimalism%20final%20for%20Mind%20-%20Guillermo%20Del%20Pinal.pdf

https://www.nature.com/articles/s41599-023-01701-2

## Look at Total Sentence Reading Time

```{r}
#look at TotalTime over trials in each experiment
pdf("charts/COMB_INT_trials_plot.pdf", width = 14, height = 14)
int_trials_plot <- ggplot(combined_int, aes(x = TrialN, y = TotalTime, color = Experiment)) +
  geom_smooth(method = "loess", alpha = .1) +
  labs(title = "Raw Sentence Reading Times Over Trials",
       y = "Reading Time (ms.)",
       x = "Trial Number") + #set up labels/title
  theme_light(base_size = 25, #default font size
              base_line_size = .5, #default line weight
              base_rect_size = 1, #default polygon border weight
              base_family = "Helvetica") + #font family Helvetica
  theme(plot.title = element_text(size = 30, #title font size
                                hjust = 0.5), #center title
      legend.background = element_rect(color = "black", linewidth = 1), #box around legend
      legend.position = "bottom") + #legend at bottom
  scale_colour_manual(values = cbPalette) #colour blind friendly palette
int_trials_plot
```

```{r}
#look at predicted TotalTime over trials in each experiment
pdf("charts/COMB_INT_trials_pred_plot.pdf", width = 14, height = 14)
int_trials_pred_plot <- ggplot(combined_int, aes(x = TrialN, y = TotalPRED, color = Experiment)) +
  geom_smooth(method = "loess", alpha = .1) +
  labs(title = "Predicted Sentence Reading Times Over Trials",
       y = "Reading Time (ms.)",
       x = "Trial Number") + #set up labels/title
  theme_light(base_size = 25, #default font size
              base_line_size = .5, #default line weight
              base_rect_size = 1, #default polygon border weight
              base_family = "Helvetica") + #font family Helvetica
  theme(plot.title = element_text(size = 30, #title font size
                                hjust = 0.5), #center title
      legend.background = element_rect(color = "black", linewidth = 1), #box around legend
      legend.position = "bottom") + #legend at bottom
  scale_colour_manual(values = cbPalette) #colour blind friendly palette
int_trials_pred_plot
```

## Look at whether you get early PP effects on early trials in EXP2 that are similar to those in EXP1

```{r}
#look at PP Reading Times over trials in each experiment
pdf("charts/COMB_INT_pp_trials_plot.pdf", width = 14, height = 14)
int_pp_trials_plot <- ggplot(combined_int, aes(x = TrialN, y = PP.RESID, color = RestrictorType)) +
  geom_smooth(method = "lm", alpha = .1) +
  labs(title = "PP Reading Times Over Trials",
       y = "Reading Time (ms.)",
       x = "Trial Number") + #set up labels/title
  theme_light(base_size = 25, #default font size
              base_line_size = .5, #default line weight
              base_rect_size = 1, #default polygon border weight
              base_family = "Helvetica") + #font family Helvetica
  theme(plot.title = element_text(size = 30, #title font size
                                hjust = 0.5), #center title
      legend.background = element_rect(color = "black", linewidth = 1), #box around legend
      legend.position = "bottom") + #legend at bottom
  facet_grid(~Experiment) +
  scale_colour_manual(values = cbPalette) #colour blind friendly palette
int_pp_trials_plot
```

```{r}
#separate trials into blocks
combined_int <- combined_int %>%
  mutate(TrialBlock = cut(TrialN, breaks = c(0, 40, 80, 120), labels = c("EARLY", "MIDDLE", "LATE")))
```

```{r}
#factorise
combined_int <- combined_int %>%
  mutate(Participant = Participant) %>%
  mutate(RestrictorType = factor(RestrictorType, levels = c("SUB", "MAX")),
         Participant = factor(Participant),
         StimID = factor(StimID),
         Experiment = factor(Experiment),
         factor(TrialBlock))
```

```{r}
#contrast code
contrasts(combined_int$RestrictorType) <- matrix(c(.5, -.5))
contrasts(combined_int$TrialBlock) <- contr.sum(3)
contrasts(combined_int$Experiment) <- matrix(c(.5, -.5))
```

```{r}
chunks_int_BLOCKS <- combined_int %>%
  pivot_longer(cols = all_of(roi_int), 
               names_to = "Chunk", 
               values_to = "ReadingTime") %>%
  group_by(RestrictorType, Chunk, Experiment, TrialBlock) %>%
  summarise(
    mean_RT = mean(ReadingTime),
    sd_RT = sd(ReadingTime),
    se_RT = sd_RT / sqrt(n()),
    .groups = "drop"
  )
```


```{r}
#plots
pdf("charts/COMB_INT_chunks_BLOCKS_plot.pdf", width = 14, height = 14)
chunks_int_BLOCKS_plot <- chunks_int_BLOCKS %>%
  ggplot(aes(x = Chunk, y = mean_RT, group = RestrictorType, color = RestrictorType)) +
  geom_ribbon(aes(ymin = mean_RT - se_RT, ymax = mean_RT + se_RT, group = RestrictorType), 
              fill = "grey", alpha = 0.2, linetype = "blank") +
  geom_line(aes(linetype = RestrictorType), linewidth = 1) +
  geom_point(aes(shape = RestrictorType), size = 3) +
  scale_y_continuous(name = "Mean Reading Time (ms.)") +
  theme_light(base_size = 20, base_family = "Helvetica") +
  scale_colour_manual(values = cbPalette) +
  facet_grid(TrialBlock~Experiment) +
  labs(title = "Reading Times (ms.): EARLY vs. MIDDLE vs. LATE Trials") +
  theme(legend.background = element_rect(color = "black", linewidth = 1),
        plot.title = element_text(hjust = 0.5, size = 25),
        strip.background = element_rect(fill = "grey90"),
        strip.text = element_text(colour = "black")) +
  scale_x_discrete(labels = c("EVERY", "NP", "PP", "VP", "SPILL"))
chunks_int_BLOCKS_plot #probably not adapted an experiment-specific reading strategy
```

```{r}
#nb: more complex models failed to converge/were singular fit
pp.int.mod1 <- lmer(PP.RESID ~ RestrictorType * Experiment * TrialBlock +
                      (1 | Participant),
                    data = combined_int)
summary(pp.int.mod1)
```

```{r}
#check model
check_model(pp.int.mod1) #ok
```

```{r}
mod_out(pp.int.mod1)
```

```{r}
int_blocks_emms <- emmeans(pp.int.mod1, ~ RestrictorType * TrialBlock | Experiment)
int_blocks_contrasts <- contrast(int_blocks_emms, interaction = "pairwise", by = "Experiment", adjust = "sidak")
```

```{r}
int_blocks_contrasts %>%
  xtable()
```

```{r}
#plot it
pdf("charts/COMB_INT_pp_emms_ixn_plot.pdf", width = 14, height = 14)
int_pp_ixn_plot <- emmip(pp.int.mod1, 
                            RestrictorType ~ Experiment,
                           CIs = TRUE,
                           CIarg = list(lwd = .5,
                                        alpha = 1)) +
  scale_colour_manual(values = cbPalette) + #colour blind friendly palette
  labs(title = "PP EMM Reading Times Interaction Plot",
       y = "Estimated Marginal Mean Reading Time (ms.)",
       x = "Experiment") + #set up labels/title
  theme_light(base_size = 25, #default font size
              base_line_size = .5, #default line weight
              base_rect_size = 1, #default polygon border weight
              base_family = "Helvetica") + #font family Helvetica
  theme(plot.title = element_text(size = 30, #title font size
                                hjust = 0.5), #center title
      legend.background = element_rect(color = "black", linewidth = 1), #box around legend
      legend.position = "bottom") #legend at bottom
int_pp_ixn_plot
```

```{r}
#plot it
pdf("charts/COMB_INT_pp_blocks_emms_ixn_plot.pdf", width = 14, height = 14)
int_pp_blocks_ixn_plot <- emmip(pp.int.mod1, 
                            RestrictorType ~ TrialBlock | Experiment,
                           CIs = TRUE,
                           CIarg = list(lwd = .5,
                                        alpha = 1)) +
  scale_colour_manual(values = cbPalette) + #colour blind friendly palette
  labs(title = "PP EMM Reading Times Interaction Plot",
       y = "Estimated Marginal Mean Reading Time (ms.)",
       x = "Experiment") + #set up labels/title
  theme_light(base_size = 25, #default font size
              base_line_size = .5, #default line weight
              base_rect_size = 1, #default polygon border weight
              base_family = "Helvetica") + #font family Helvetica
  theme(plot.title = element_text(size = 30, #title font size
                                hjust = 0.5), #center title
      legend.background = element_rect(color = "black", linewidth = 1), #box around legend
      legend.position = "bottom") #legend at bottom
int_pp_blocks_ixn_plot
```