---
title: "Combined Analyses"
output: html_document
---

# Set Up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 14)
```

```{r}
#clear workspace
rm(list=ls()) 
```

```{r}
#colour blind friendly colour palette for dataviz.
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
#custom functions

#get model output as latex
mod_out <- function(model_name) {
  # Ensure the model is of class 'lmerMod'
  class(model_name) <- "lmerMod"
  
  # Generate LaTeX output with stargazer
  stargazer(model_name,
            out = "model_output.tex", 
            star.cutoffs = c(0.05, 0.01, 0.001), 
            digits = 2, 
            no.space = TRUE, 
            omit.stat = "LL", 
            single.row = TRUE, 
            align = TRUE, 
            header = FALSE, 
            intercept.bottom = FALSE)
}

#output summary stats as latex
sum_tbl <- function(combined_latin) {
  summary_table <- combined_latin %>%
    group_by(Condition) %>%
    summarise(
      Mean_PP = mean(PP.RESID, na.rm = TRUE),
      SD_PP = sd(PP.RESID, na.rm = TRUE),
      Mean_VP = mean(VP.RESID, na.rm = TRUE),
      SD_VP = sd(VP.RESID, na.rm = TRUE)
    ) %>%
    ungroup() %>%
    mutate(across(where(is.numeric), ~ round(.x, 2))) %>%
    kable(format = "latex", booktabs = TRUE,
          col.names = c("Condition", "Mean (ms)", "SD (ms)", 
                        "Mean (ms)", "SD (ms)")) %>%
    kable_styling(latex_options = c("striped", "hold_position")) %>%
    add_header_above(c(" " = 1, "PP Region" = 2, "VP Region" = 2))
  
  cat(summary_table)
}
```


```{r, message=FALSE}
#load packages - nb: not all packages used
library(tidyverse)      #for tidying and wrangling
library(lme4)           #for LMM building
library(lmerTest)       #for better output from LMMs
library(emmeans)        #for our planned comparisons
library(performance)    #check model assumptions
library(buildmer)       #assists with iterating through covariates and ixns
library(ggpubr)         #plotting
library(texreg)         #for LaTeX table output
library(kableExtra)     #table formatting
library(xtable)         #table outputs
library(stargazer)      #lmm output
library(grid)           #multiple plots
library(patchwork)      #for plotting
library(stringr)        #string manipulation
library(scales)         #aesthetics
library(ggthemes)       #aesthetics
library(fitdistrplus)   #to check distribution
library(car)            #for vif() function
library(Hmisc)          #for rcorr() function
```

```{r}
#control scientific notation - makes reading p-values easier
options("scipen"=100000, "digits"=10)
#increase no. of calcs for emmeans functions
emm_options(pbkrtest.limit = 10000,
            lmerTest.limit = 10000)
```

# EXP2/3

```{r, message=FALSE}
#load datasets
exp2_tidied <- read_csv("exp2_tidied.csv")
exp3_tidied <- read_csv("exp3_tidied.csv")
```

## Wrangling

```{r}
#add experiment ID
exp3_tidied <- exp3_tidied %>%
  mutate(Experiment = "EXP3")
```

```{r}
#add experiment ID
exp2_tidied <- exp2_tidied %>%
  mutate(Experiment = "EXP2") %>%
  mutate(Participant = Participant + 59) #add 59 to participant number so we have unique participants
```

```{r}
#find columns in common
common_cols_latin <- Reduce(intersect, list(colnames(exp2_tidied), colnames(exp3_tidied)))
```

```{r}
#combine
combined_latin <- bind_rows(
  exp2_tidied %>% dplyr::select(all_of(common_cols_latin)),
  exp3_tidied %>% dplyr::select(all_of(common_cols_latin))
)
```

```{r}
#count participants - 119 (=60+59)
combined_latin %>%
  distinct(Participant) %>%
  count()
```

```{r}
#count conditions
combined_latin %>%
  count(Condition)
```

```{r}
#uneven number of obs - downsample
set.seed(76543) #ensure reproducibility
combined_latin <- combined_latin %>%
  group_by(Condition) %>%
  sample_n(885) %>%
  ungroup()
```

```{r}
#count conditions
combined_latin %>%
  count(Condition)
```

```{r}
#factorise
combined_latin <- combined_latin %>%
  mutate(RestrictorType = factor(RestrictorType, levels = c("OTH", "SUB", "MAX")),
       RestrictorPosition = factor(RestrictorPosition, levels = c("EXT", "INT")),
       Participant = factor(Participant),
       StimID = factor(StimID))
```

```{r}
#contrast code
contrasts(combined_latin$RestrictorPosition) <- matrix(c(.5, -.5))
contrasts(combined_latin$RestrictorType) <- contr.sum(3)
```

## Descriptives & dataviz for RTs for each chunk

```{r}
#output summary stats
sum_tbl(combined_latin)
```

```{r, message=FALSE, warning=FALSE}
#pull out chunk names
roi_latin <- grep("^C\\d+\\.RT.RESID$|^SPILL\\.RT.RESID$", names(combined_latin), value = TRUE)

chunks_latin <- combined_latin %>%
  pivot_longer(cols = all_of(roi_latin), 
               names_to = "Chunk", 
               values_to = "ReadingTime") %>%
  group_by(Condition, Chunk) %>%
  summarise(
    mean_RT = mean(ReadingTime),
    sd_RT = sd(ReadingTime),
    se_RT = sd_RT / sqrt(n())
  ) %>%
  separate(Condition, into = c("RestrictorPosition", "RestrictorType")) %>%
  mutate(ChunkLabel = case_when(
    RestrictorPosition == "EXT" & Chunk == "C1.RT.RESID" ~ "EVERY",
    RestrictorPosition == "EXT" & Chunk == "C2.RT.RESID" ~ "NP",
    RestrictorPosition == "EXT" & Chunk == "C3.RT.RESID" ~ "VP",
    RestrictorPosition == "EXT" & Chunk == "C4.RT.RESID" ~ "PP",
    RestrictorPosition == "EXT" & Chunk == "SPILL.RT.RESID" ~ "SPILL",
    RestrictorPosition == "INT" & Chunk == "C1.RT.RESID" ~ "EVERY",
    RestrictorPosition == "INT" & Chunk == "C2.RT.RESID" ~ "NP",
    RestrictorPosition == "INT" & Chunk == "C3.RT.RESID" ~ "PP",
    RestrictorPosition == "INT" & Chunk == "C4.RT.RESID" ~ "VP",
    RestrictorPosition == "INT" & Chunk == "SPILL.RT.RESID" ~ "SPILL",
    TRUE ~ Chunk
  ))

```

```{r}
#plots
latin_ext <- chunks_latin %>%
  filter(RestrictorPosition == "EXT") %>%
  ggplot(aes(x = Chunk, y = mean_RT, group = RestrictorType, color = RestrictorType)) +
  geom_ribbon(aes(ymin = mean_RT - se_RT, ymax = mean_RT + se_RT, group = RestrictorType), 
              fill = "grey", alpha = 0.2, linetype = "blank") +
  geom_line(aes(linetype = RestrictorType), linewidth = 1) +  
  geom_point(aes(shape = RestrictorType), size = 3) +  
  scale_y_continuous(name = "Mean Reading Time (ms.)") +
  theme_light(base_size = 20, base_family = "Helvetica") +
  scale_colour_manual(values = cbPalette) +  
  facet_grid(vars(RestrictorPosition), switch = "y") +
  theme(legend.background = element_rect(color = "black", linewidth = 1),
        plot.title = element_text(hjust = 0.5, size = 25),
        strip.background = element_rect(fill = "grey90"),
        strip.text = element_text(colour = "black")) +
  scale_x_discrete(labels = c("EVERY", "NP", "VP", "PP", "SPILL"))

latin_int <- chunks_latin %>%
  filter(RestrictorPosition == "INT") %>%
  ggplot(aes(x = Chunk, y = mean_RT, group = RestrictorType, color = RestrictorType)) +
  geom_ribbon(aes(ymin = mean_RT - se_RT, ymax = mean_RT + se_RT, group = RestrictorType), 
              fill = "grey", alpha = 0.2, linetype = "blank") +
  geom_line(aes(linetype = RestrictorType), linewidth = 1) +  
  geom_point(aes(shape = RestrictorType), size = 3) +  
  scale_y_continuous(name = "Mean Reading Time (ms.)") +
  theme_light(base_size = 20, base_family = "Helvetica") +
  scale_colour_manual(values = cbPalette) +  
  facet_grid(vars(RestrictorPosition), switch = "y") +
  theme(legend.background = element_rect(color = "black", linewidth = 1),
        plot.title = element_text(hjust = 0.5, size = 25),
        strip.background = element_rect(fill = "grey90"),
        strip.text = element_text(colour = "black")) +
  scale_x_discrete(labels = c("EVERY", "NP", "PP", "VP", "SPILL"))
```

```{r}
#combine plots
pdf("charts/COMB_LATIN_chunks.pdf", width = 14, height = 14)
latin_ext + 
  latin_int + 
  plot_layout(ncol = 1,
              guides = "collect",
              axis_titles = "collect") +
  plot_annotation(title = "Mean Chunk Reading Times (ms.)",
                  theme = theme(plot.title = element_text(hjust = 0.5,
                                                          size = 30)))
```

## Reading Time Analysis for Preposition Phrase

```{r}
#dataviz for condition - logged for better viz.
pdf("charts/COMB_LATIN_pp_plot.pdf", width = 14, height = 14)
pp_ltn <- combined_latin %>%
  ggplot(aes(x = RestrictorType, #group by condition on x axis
             y = PP.RESID.LOG, #log RTs on y axis
             fill = RestrictorType)) + #colour by condition
  theme_light(base_size = 25, #default font size
              base_line_size = .5, #default line weight
              base_rect_size = 1, #default polygon border weight
              base_family = "Helvetica") + #font family Helvetica
  geom_violin(trim = FALSE) + #violin, don't trim ends
  geom_boxplot(width = 0.3, #boxplot
               outlier.shape = NA) + #don't show outlier dots
  scale_fill_manual(values = cbPalette) + #colour blind friendly palette
  labs(title = "PP Reading Times (ms.)",
       y = "Logged Reading Time (ms.)",
       x = NULL) +
  theme(legend.background = element_rect(color = "black", linewidth = 1), #box around legend
        plot.title = element_text(hjust = 0.5, size = 30), #center title, title font size
        legend.position = "bottom") + #legend at bottom
  coord_flip() + #flip coordinates so it's horizontal
  facet_wrap(~RestrictorPosition) +
  theme(legend.background = element_rect(color = "black", linewidth = 1),
        plot.title = element_text(hjust = 0.5, size = 25),
        strip.background = element_rect(fill = "grey90"),
        strip.text = element_text(colour = "black"))
pp_ltn
```

### Modelling

```{r, results=FALSE, message=FALSE}
#building a buildmer model
pp.ltn.blmr1 <- buildmer(PP.RESID ~ RestrictorType * RestrictorPosition +
                       (1 + RestrictorType * RestrictorPosition | Participant) +
                       (1 + RestrictorType * RestrictorPosition | StimID),
                     buildmerControl = buildmerControl(include = ~RestrictorType * RestrictorPosition),
                     data = combined_latin)
```

```{r}
#model summary
summary(pp.ltn.blmr1)
```

```{r}
#increase ranfx
pp.ltn1 <- lmer(PP.RESID ~ RestrictorType * RestrictorPosition +
                  (1 + RestrictorPosition | Participant) +
                  (1 | StimID),
                data = combined_latin)
summary(pp.ltn1) #OK, main fx, ixn
```

```{r}
#increase ranfx complexity
pp.ltn2 <- lmer(PP.RESID ~ RestrictorType * RestrictorPosition +
                  (1 + RestrictorPosition | Participant) +
                  (1 + RestrictorPosition | StimID),
                data = combined_latin)
summary(pp.ltn2) #singular fit
```

```{r}
#increase ranfx complexity
pp.ltn3 <- lmer(PP.RESID ~ RestrictorType * RestrictorPosition +
                  (1 + RestrictorType + RestrictorPosition | Participant) +
                  (1 | StimID),
                data = combined_latin)
summary(pp.ltn3) #singular fit, convergence issue
```

```{r}
#check model
check_model(pp.ltn1) #OK
```

```{r}
#add covars
pp.ltn4 <- lmer(PP.RESID ~ RestrictorType * RestrictorPosition + 
                  TrialN + Experiment +
                  (1 + RestrictorPosition | Participant) +
                  (1 | StimID),
                data = combined_latin)
summary(pp.ltn4) #ok, main fx, ixn
```

```{r}
check_model(pp.ltn4) #ok
```

```{r}
#compare - covar mod better
anova(pp.ltn1, pp.ltn4)
```

```{r}
#model output
mod_out(pp.ltn4)
```

```{r}
#emms
pp.ltn_emms1 <- emmeans(pp.ltn4, ~ RestrictorType | RestrictorPosition,
                   adjust = "none") #we'll adjust when doing contrasts
```

```{r}
#contrasts
pp.ltn_contrasts1 <- contrast(pp.ltn_emms1,
                              adjust = "tukey",
                              "pairwise",
                              simple = "RestrictorType")
```

```{r}
#print latex
pp.ltn_contrasts1 %>%
  xtable() 
#early fx could be priming, later fx less so.
#in future work, I'll try and devise
#a better paradigm to mitigate against this
```
```{r}
#plot it
pdf("charts/COMB_LATIN_emms_ixn_plot.pdf", width = 14, height = 14)
ltn_pp_ixn_plot <- emmip(pp.ltn4, 
                            RestrictorType ~ RestrictorPosition,
                           CIs = TRUE,
                           CIarg = list(lwd = .5,
                                        alpha = 1)) +
  scale_colour_manual(values = cbPalette) + #colour blind friendly palette
  labs(title = "PP EMM Reading Times Interaction Plot",
       y = "Estimated Marginal Mean Reading Time (ms.)",
       x = "RestrictorPosition") + #set up labels/title
  theme_light(base_size = 25, #default font size
              base_line_size = .5, #default line weight
              base_rect_size = 1, #default polygon border weight
              base_family = "Helvetica") + #font family Helvetica
  theme(plot.title = element_text(size = 30, #title font size
                                hjust = 0.5), #center title
      legend.background = element_rect(color = "black", linewidth = 1), #box around legend
      legend.position = "bottom") #legend at bottom
ltn_pp_ixn_plot
```

## Reading Time Analysis for Verb Phrase

```{r}
#dataviz for condition - logged for better viz.
pdf("charts/COMB_LATIN_vp_plot.pdf", height = 14, width = 14)
vp_viz <- combined_latin %>%
  ggplot(aes(x = RestrictorType, #group by condition on x axis
             y = VP.RESID.LOG, #log RTs on y axis
             fill = RestrictorType)) + #colour by condition
  theme_light(base_size = 25, #default font size
              base_line_size = .5, #default line weight
              base_rect_size = 1, #default polygon border weight
              base_family = "Helvetica") + #font family Helvetica
  geom_violin(trim = FALSE) + #violin, don't trim ends
  geom_boxplot(width = 0.3, #boxplot
               outlier.shape = NA) + #don't show outlier dots
  scale_fill_manual(values = cbPalette) + #colour blind friendly palette
  labs(title = "VP Reading Times (ms.)",
       y = "Logged Reading Time (ms.)",
       x = NULL) +
  theme(legend.background = element_rect(color = "black", linewidth = 1), #box around legend
        plot.title = element_text(hjust = 0.5, size = 30), #center title, title font size
        legend.position = "bottom") + #legend at bottom
  coord_flip() + #flip coordinates so it's horizontal
  facet_wrap(~RestrictorPosition) +
  theme(legend.background = element_rect(color = "black", linewidth = 1),
        plot.title = element_text(hjust = 0.5, size = 25),
        strip.background = element_rect(fill = "grey90"),
        strip.text = element_text(colour = "black"))
#print chart
vp_viz
```

### Modelling

```{r}
#buildmer model
vp.ltn.blmr <- buildmer(VP.RESID ~ RestrictorType * RestrictorPosition +
                       (1 + RestrictorType * RestrictorPosition | Participant) +
                       (1 + RestrictorType * RestrictorPosition | StimID),
                       data = combined_latin)
```

```{r}
summary(vp.ltn.blmr) #main fx, ixn
```

```{r}
#manual model
vp.ltn1 <- lmer(VP.RESID ~ RestrictorType * RestrictorPosition +
                  (1 + RestrictorPosition | Participant) +
                  (1 | StimID),
                data = combined_latin)
summary(vp.ltn1) #main fx, ixn
```

```{r}
#increase ranfx complexity
vp.ltn2 <- lmer(VP.RESID ~ RestrictorType * RestrictorPosition +
                  (1 + RestrictorPosition | Participant) +
                  (1 + RestrictorType | StimID),
                data = combined_latin)
summary(vp.ltn2) #singular fit
```

```{r}
#increase ranfx complexity
vp.ltn3 <- lmer(VP.RESID ~ RestrictorType * RestrictorPosition +
                  (1 + RestrictorPosition | Participant) +
                  (1 + RestrictorPosition | StimID),
                data = combined_latin)
summary(vp.ltn3) #singular fit
```

```{r}
#add covars
vp.ltn4 <- lmer(VP.RESID ~ RestrictorType * RestrictorPosition + scale(PP.RT) +
                  TrialN + Experiment +
                  (1 + RestrictorPosition | Participant) +
                  (1 | StimID),
                data = combined_latin)
summary(vp.ltn4) #main fx, ixn
```

```{r}
#check model
check_model(vp.ltn4) #ok
```

```{r}
#compare - covar mod better
anova(vp.ltn1, vp.ltn4)
```

```{r}
#model output
mod_out(vp.ltn4)
```

```{r}
#emms
vp.ltn_emms1 <- emmeans(vp.ltn4, ~ RestrictorType * RestrictorPosition,
                   adjust = "none")
vp.ltn_emms1
```

```{r}
#just look at INT
vp.ltn.int_emms1 <- subset(vp.ltn_emms1, RestrictorPosition == "INT")
```

```{r}
#contrasts
contrast(vp.ltn.int_emms1,
         adjust = "tukey",
         "pairwise") %>%
  xtable()
```

```{r}
#checking that INT conditions are slower that EXT conditions overall
vp.ltn.pos.emms1 <- emmeans(vp.ltn4,
                       pairwise ~ RestrictorPosition,
                       adjust = "none") #only one comparison made
vp.ltn.pos.emms1
```

```{r}
#plot it
pdf("charts/COMB_LATIN_vp_ixn_plot.pdf", height = 14, width = 14)
vp_ixn_plot <- emmip(vp.ltn4,
                     RestrictorType ~ RestrictorPosition,
                     CIs = TRUE,
                     CIarg = list(lwd = .5,
                                  alpha = 1)) +
  scale_colour_manual(values = cbPalette) + #colour blind friendly palette
  labs(title = "VP EMM Reading Times Interaction Plot",
       y = "Estimated Marginal Mean Reading Time (ms.)",
       x = "RestrictorPosition") + #set up labels/title
  theme_light(base_size = 25, #default font size
              base_line_size = .5, #default line weight
              base_rect_size = 1, #default polygon border weight
              base_family = "Helvetica") + #font family Helvetica
  theme(plot.title = element_text(size = 30, #title font size
                                  hjust = 0.5), #center title
        legend.background = element_rect(color = "black", linewidth = 1), #box around legend
        legend.position = "bottom") #legend at bottom
vp_ixn_plot
```

### Could there be priming effects?

```{r}
#add cumulative reading time
combined_latin_prime <- combined_latin %>%
  filter(Experiment == "EXP3") %>%
  group_by(Participant, TrialN) %>%
  mutate(CumulativeRT = case_when(RestrictorPosition == "INT" ~ sum(C1.RT, C2.RT), #every + N
                                  RestrictorPosition == "EXT" ~ sum(C1.RT, C2.RT, C3.RT))) %>% #every + N + PP
  mutate(RestrictorType = factor(RestrictorType)) #refactorise - now just two levels
```

```{r}
#apply contrasts
contrasts(combined_latin_prime$RestrictorType) <- matrix(c(.5, -.5))
```

```{r}
pdf("charts/COMB_LATIN_cumulativeRT_effect.pdf", width = 14, height = 14)
ltn_cumulative_plot_int <- combined_latin_prime %>%
  ggplot(aes(x = log(CumulativeRT), y = PP.RESID, color = RestrictorType, linetype = RestrictorType)) +
  geom_smooth(method = "loess", alpha = 0.1) +
  labs(x = "Cumulative Reading Time", y = "PP Reading Time") +
  theme_light(base_size = 25) +
  facet_wrap(~RestrictorPosition, switch = "y") +
  theme(plot.title = element_text(size = 30, hjust = 0.5),
        legend.background = element_rect(color = "black", linewidth = 1),
        legend.position = "bottom") +
  scale_colour_manual(values = cbPalette)
ltn_cumulative_plot_int
```

```{r}
#model
int.prime.blmr1 <- buildmer(PP.RESID ~ log(CumulativeRT) * RestrictorType * RestrictorPosition +
                               TrialN +
                               (1 + CumulativeRT * RestrictorType * RestrictorPosition| Participant) +
                               (1 + CumulativeRT * RestrictorType * RestrictorPosition | StimID),
                       data = combined_latin_prime)
summary(int.prime.blmr1)
```

```{r}
#model
int.prime.mod1 <- lmer(PP.RESID ~ log(CumulativeRT) * RestrictorType * RestrictorPosition +
                               TrialN +
                               (1 + RestrictorPosition| Participant),
                       data = combined_latin_prime)
summary(int.prime.mod1)
```

```{r}
mod_out(int.prime.mod1)
```

```{r}
combined_latin_prime$PredictedSlope <- predict(int.prime.mod1)
```

```{r}
pdf("charts/COMB_LATIN_cumulativeRT_predicted_effect.pdf", width = 14, height = 14)
combined_latin_prime_predicted_effect <- combined_latin_prime %>%
  ggplot(aes(x = CumulativeRT, y = PredictedSlope, color = RestrictorType, linetype = RestrictorType)) +
  geom_smooth(method = "loess", se = TRUE, alpha = 0.1) +
  labs(x = "Cumulative Reading Time", y = "PP Reading Time") +
  theme_light(base_size = 25) +
  facet_wrap(~RestrictorPosition) +
  theme(plot.title = element_text(size = 30, hjust = 0.5),
        legend.background = element_rect(color = "black", linewidth = 1),
        legend.position = "bottom") +
  scale_colour_manual(values = cbPalette)
combined_latin_prime_predicted_effect
```

# EXP1/2 INT-SUB vs. INT-MAX

## Wrangling

```{r}
#load exp1 dataset
exp1_tidied <- read_csv("exp1_tidied.csv") %>%
  mutate(Experiment = "EXP1") %>%
  filter(Condition != "NIL") %>%
  mutate(has_comp = as.character(has_comp))
```

```{r}
exp2_COPY_tidied <- exp2_tidied %>%
  mutate(has_comp = as.character(has_comp)) %>% #change has_comp to dbl
  mutate(Participant = Participant + 1) #add 1 to avoid overlap
```

```{r}
#find columns in common
common_cols_int <- Reduce(intersect, list(colnames(exp1_tidied), colnames(exp2_COPY_tidied)))
```

```{r}
#combine
combined_int <- bind_rows(
  exp1_tidied %>% dplyr::select(all_of(common_cols_int)),
  exp2_COPY_tidied %>%
    filter(RestrictorPosition == "INT")
  %>% dplyr::select(all_of(common_cols_int)))
```

```{r}
#count participants - 120
combined_int %>%
  distinct(Participant) %>%
  count()
```
```{r}
#ensure Condition is labelled consistently
combined_int <- combined_int %>%
  mutate(Condition = RestrictorType)
```

```{r}
#add Total Sentence Reading time raw and predicted
combined_int <- combined_int %>%
  rowwise() %>%
  mutate(TotalTime = sum(C1.RT, C2.RT, C3.RT, C4.RT, SPILL.RT),
         TotalPRED = sum(C1.PRED, C2.PRED, C3.PRED, C4.PRED, SPILL.PRED)) %>%
  ungroup()
```

## Descriptives & dataviz for RTs for each chunk

```{r, message=FALSE, warning=FALSE}
#pull out chunk names
roi_int <- grep("^C\\d+\\.RT.RESID$|^SPILL\\.RT.RESID$", names(combined_int), value = TRUE)

chunks_int <- combined_int %>%
  pivot_longer(cols = all_of(roi_int), 
               names_to = "Chunk", 
               values_to = "ReadingTime") %>%
  group_by(RestrictorType, Chunk, Experiment) %>%
  summarise(
    mean_RT = mean(ReadingTime),
    sd_RT = sd(ReadingTime),
    se_RT = sd_RT / sqrt(n())
  )
```

```{r}
#plots
pdf("charts/COMB_INT_chunks_plot.pdf", width = 14, height = 14)
combined_int_chunks_plot <- chunks_int %>%
  ggplot(aes(x = Chunk, y = mean_RT, group = RestrictorType, color = RestrictorType)) +
  geom_ribbon(aes(ymin = mean_RT - se_RT, ymax = mean_RT + se_RT, group = RestrictorType), 
              fill = "grey", alpha = 0.2, linetype = "blank") +
  geom_line(aes(linetype = RestrictorType), linewidth = 1) +
  geom_point(aes(shape = RestrictorType), size = 3) +
  scale_y_continuous(name = "Mean Reading Time (ms.)") +
  theme_light(base_size = 20, base_family = "Helvetica") +
  scale_colour_manual(values = cbPalette) +
  facet_grid(~ Experiment) +
  theme(legend.background = element_rect(color = "black", linewidth = 1),
        plot.title = element_text(hjust = 0.5, size = 25),
        strip.background = element_rect(fill = "grey90"),
        strip.text = element_text(colour = "black")) +
  scale_x_discrete(labels = c("EVERY", "NP", "PP", "VP", "SPILL"))
combined_int_chunks_plot
```

## Look at Total Sentence Reading Time

```{r}
combined_int %>%
  group_by(Experiment) %>%
  summarise("Mean Raw Reading Time" = mean(TotalTime),
            "SD Raw Reading Time" = sd(TotalTime),
            "Mean Predicted Reading Time" = mean(TotalPRED),
            "SD Predicted Reading Time" = sd(TotalPRED)) %>%
  xtable()
```

```{r}
#look at TotalTime over trials in each experiment
pdf("charts/COMB_INT_trials_plot.pdf", width = 14, height = 14)
int_trials_plot <- ggplot(combined_int, aes(x = TrialN, y = TotalTime, color = Experiment)) +
  geom_smooth(method = "gam", alpha = .1) +
  labs(title = "Raw Sentence Reading Times Over Trials",
       y = "Reading Time (ms.)",
       x = "Trial Number") + #set up labels/title
  theme_light(base_size = 25, #default font size
              base_line_size = .5, #default line weight
              base_rect_size = 1, #default polygon border weight
              base_family = "Helvetica") + #font family Helvetica
  theme(plot.title = element_text(size = 30, #title font size
                                hjust = 0.5), #center title
      legend.background = element_rect(color = "black", linewidth = 1), #box around legend
      legend.position = "bottom") + #legend at bottom
  scale_colour_manual(values = cbPalette) #colour blind friendly palette
int_trials_plot
```

```{r}
#look at predicted TotalTime over trials in each experiment
pdf("charts/COMB_INT_trials_pred_plot.pdf", width = 14, height = 14)
int_trials_pred_plot <- ggplot(combined_int, aes(x = TrialN, y = TotalPRED, color = Experiment)) +
  geom_smooth(method = "lm", alpha = .1) +
  labs(title = "Predicted Sentence Reading Times Over Trials",
       y = "Reading Time (ms.)",
       x = "Trial Number") + #set up labels/title
  theme_light(base_size = 25, #default font size
              base_line_size = .5, #default line weight
              base_rect_size = 1, #default polygon border weight
              base_family = "Helvetica") + #font family Helvetica
  theme(plot.title = element_text(size = 30, #title font size
                                hjust = 0.5), #center title
      legend.background = element_rect(color = "black", linewidth = 1), #box around legend
      legend.position = "bottom") + #legend at bottom
  scale_colour_manual(values = cbPalette) #colour blind friendly palette
int_trials_pred_plot
```

```{r}
#factorise
combined_int <- combined_int %>%
  mutate(Participant = Participant) %>%
  mutate(RestrictorType = factor(RestrictorType, levels = c("SUB", "MAX")),
         Participant = factor(Participant),
         StimID = factor(StimID))
```

```{r}
#factorise experiment
combined_int <- combined_int %>%
  mutate(Experiment = factor(Experiment))
```

```{r}
#contrast code
contrasts(combined_int$RestrictorType) <- matrix(c(.5, -.5))
contrasts(combined_int$Experiment) <- matrix(c(.5, -.5))
```

```{r}
#model
TotalTime.mod1 <- lmer(TotalTime ~ Experiment * scale(TotalPRED) + scale(TrialN) +
                             (1 | Participant),
                           data = combined_int)
summary(TotalTime.mod1)
```

```{r}
mod_out(TotalTime.mod1)
```

```{r}
combined_int$Predicted <- predict(TotalTime.mod1)
```

## Look at whether you get early PP effects on early trials in EXP2 that are similar to those in EXP1

```{r}
#separate trials into blocks
combined_int <- combined_int %>%
  mutate(TrialBlock = cut(TrialN, breaks = c(0, 40, 80, 120), labels = c("EARLY", "MIDDLE", "LATE"))) %>%
  mutate(TrialBlock = factor(TrialBlock))
```

```{r}
#contrast code
contrasts(combined_int$TrialBlock) <- contr.sum(3)
```

```{r}
combined_int %>%
  group_by(Experiment, TrialBlock, RestrictorType) %>%
  summarise("Mean Reading Time" = mean(PP.RESID),
                             "SD Reading Time" = sd(PP.RESID)) %>%
  xtable()
```


```{r}
#look at PP Reading Times over trials in each experiment
pdf("charts/COMB_INT_pp_trials_plot.pdf", width = 14, height = 14)
int_pp_trials_plot <- ggplot(combined_int, aes(x = TrialN, y = PP.RESID, color = RestrictorType)) +
  geom_smooth(method = "lm", alpha = .1) +
  labs(title = "PP Reading Times Over Trials",
       y = "Reading Time (ms.)",
       x = "Trial Number") + #set up labels/title
  theme_light(base_size = 25, #default font size
              base_line_size = .5, #default line weight
              base_rect_size = 1, #default polygon border weight
              base_family = "Helvetica") + #font family Helvetica
  theme(plot.title = element_text(size = 30, #title font size
                                hjust = 0.5), #center title
      legend.background = element_rect(color = "black", linewidth = 1), #box around legend
      legend.position = "bottom") + #legend at bottom
  facet_grid(~Experiment) +
  scale_colour_manual(values = cbPalette) #colour blind friendly palette
int_pp_trials_plot
```

```{r}
chunks_int_BLOCKS <- combined_int %>%
  pivot_longer(cols = all_of(roi_int), 
               names_to = "Chunk", 
               values_to = "ReadingTime") %>%
  group_by(RestrictorType, Chunk, Experiment, TrialBlock) %>%
  summarise(
    mean_RT = mean(ReadingTime),
    sd_RT = sd(ReadingTime),
    se_RT = sd_RT / sqrt(n()),
    .groups = "drop"
  )
```

```{r}
#plots
pdf("charts/COMB_INT_chunks_BLOCKS_plot.pdf", width = 14, height = 14)
chunks_int_BLOCKS_plot <- chunks_int_BLOCKS %>%
  ggplot(aes(x = Chunk, y = mean_RT, group = RestrictorType, color = RestrictorType)) +
  geom_ribbon(aes(ymin = mean_RT - se_RT, ymax = mean_RT + se_RT, group = RestrictorType), 
              fill = "grey", alpha = 0.2, linetype = "blank") +
  geom_line(aes(linetype = RestrictorType), linewidth = 1) +
  geom_point(aes(shape = RestrictorType), size = 3) +
  scale_y_continuous(name = "Mean Reading Time (ms.)") +
  theme_light(base_size = 20, base_family = "Helvetica") +
  scale_colour_manual(values = cbPalette) +
  facet_grid(TrialBlock~Experiment) +
  labs(title = "Reading Times (ms.): EARLY vs. MIDDLE vs. LATE Trials") +
  theme(legend.background = element_rect(color = "black", linewidth = 1),
        plot.title = element_text(hjust = 0.5, size = 25),
        strip.background = element_rect(fill = "grey90"),
        strip.text = element_text(colour = "black")) +
  scale_x_discrete(labels = c("EVERY", "NP", "PP", "VP", "SPILL"))
chunks_int_BLOCKS_plot #probably not adapted an experiment-specific reading strategy
```

```{r}
#nb: more complex models failed to converge/were singular fit
pp.int.mod1 <- lmer(PP.RESID ~ RestrictorType * Experiment * TrialBlock +
                      (1 | Participant),
                    data = combined_int)
summary(pp.int.mod1)
```

```{r}
#check model
check_model(pp.int.mod1) #ok
```

```{r}
mod_out(pp.int.mod1)
```

```{r}
int_blocks_emms <- emmeans(pp.int.mod1, ~ Experiment * TrialBlock * RestrictorType)
int_blocks_contrasts <- contrast(int_blocks_emms, interaction = "pairwise", by = "Experiment", adjust = "sidak")
```

```{r}
int_blocks_contrasts %>%
  xtable()
```

```{r}
#plot it
pdf("charts/COMB_INT_pp_emms_ixn_plot.pdf", width = 14, height = 14)
int_pp_ixn_plot <- emmip(pp.int.mod1, 
                            RestrictorType ~ Experiment,
                           CIs = TRUE,
                           CIarg = list(lwd = .5,
                                        alpha = 1)) +
  scale_colour_manual(values = cbPalette) + #colour blind friendly palette
  labs(title = "PP EMM Reading Times Interaction Plot",
       y = "Estimated Marginal Mean Reading Time (ms.)",
       x = "Experiment") + #set up labels/title
  theme_light(base_size = 25, #default font size
              base_line_size = .5, #default line weight
              base_rect_size = 1, #default polygon border weight
              base_family = "Helvetica") + #font family Helvetica
  theme(plot.title = element_text(size = 30, #title font size
                                hjust = 0.5), #center title
      legend.background = element_rect(color = "black", linewidth = 1), #box around legend
      legend.position = "bottom") #legend at bottom
int_pp_ixn_plot
```

```{r}
#plot it
pdf("charts/COMB_INT_pp_blocks_emms_ixn_plot.pdf", width = 14, height = 14)
int_pp_blocks_ixn_plot <- emmip(pp.int.mod1, 
                            RestrictorType ~ TrialBlock | Experiment,
                           CIs = TRUE,
                           CIarg = list(lwd = .5,
                                        alpha = 1)) +
  scale_colour_manual(values = cbPalette) + #colour blind friendly palette
  labs(title = "PP EMM Reading Times Interaction Plot",
       y = "Estimated Marginal Mean Reading Time (ms.)",
       x = "Experiment") + #set up labels/title
  theme_light(base_size = 25, #default font size
              base_line_size = .5, #default line weight
              base_rect_size = 1, #default polygon border weight
              base_family = "Helvetica") + #font family Helvetica
  theme(plot.title = element_text(size = 30, #title font size
                                hjust = 0.5), #center title
      legend.background = element_rect(color = "black", linewidth = 1), #box around legend
      legend.position = "bottom") #legend at bottom
int_pp_blocks_ixn_plot
```